This chapter aims to provide the reader with an overview of current research and state-of-the-art in the field of Cellular V2X communication and Cooperative Perception.
\par
\bigskip

In accordance with the multiple goals of this thesis introduced in \autoref{ch:introduction} and \autoref{ch:problem_description}, related work can, in the broadest sense, be separated into three sections. On the one hand, relevant publications about \textbf{environment modeling, traffic scene representation and message exchange formats} for Cooperative Perception are examined in \autoref{sec:related_work:environment_modeling_state_representation}. Secondly, an overview of existing \textbf{Cooperative Perception systems} is presented and different approaches are discussed in view of this work in \autoref{sec:related_work:cooperative_perception}. Finally, related work on \textbf{cellular-based V2X} is presented.

\section{Environment Modeling \& State Representation}
\label{sec:related_work:environment_modeling_state_representation}
An essential requirement for high-level CP is to have a uniform way to first model the current environment state and second represent that information. It is worth noting that this does not necessarily hold true for low- or feature level CP (see \autoref{subsec:background:sensor_fusion}), where either raw sensor readings or only basic information is shared.

An appropriate state representation should, at a minimum, include information about position and dynamics of the sender vehicle as well as of all other surrounding traffic participants. For the former, the European Telecommunications Standards Institute (ETSI) has defined a standard for so called \textbf{Cooperative Awareness Messages} (CAM) \cite{EuropeanTelecommunicationsStandardsInstituteETSI2011}, which is, among others, used by \cite{Rauch2011}. It includes, among others, the sender vehicle's type, its dimensions, position, heading, speed and acceleration as well as respective confidences.

To share information about surrounding obstacles in addition, the simplest way is to send \textbf{object lists} that include such. For this purpose, \cite{Rauch2011} defined the \textbf{Cooperative Perception Message} (CPM) in 2011. Since 2017, the ETSI is working on a similar specification with the same name \cite{EuropeanTelecommunicationsStandardsInstituteETSI2019}. It is worth noting that CAMs and CPMs define both \textit{model} and \textit{representation format} at once. The latter includes an object list of up to 255 traffic participants \cite{Thandavarayan2019} with attributes similar to those included in CAMs. However, none of both specifications is publicly accessible, yet. Therefore, they can not serve as a basis for this thesis.

Despite pure object lists, a way to share the perception of one's local environment in addition is to model it in the form of \textbf{occupancy grids} or \textit{driveable area} maps \cite{pieringermodellierung}. Because, to the best of our knowledge, no CP solution exists that relies on exchanging occupancy grids, a reference on how to best \textit{represent} them for CP can not be provided.

\cite{Stiller2012} follows a different approach as the authors show a way to represent the current local world state as the instantiation of a \textbf{Markov Logic Network}, in which weights represent uncertainty about the true state of an observation. This representation is inherently graphical and by incorporating first-order logic, the underlying model also allows for basic inference, in theory. A specification of what objects and relations to include to comprehensively model a traffic scene is not provided.

\cite{Kohlhaas2014} first introduced the concept of \textbf{semantic scene representation} (or \textit{semantic state representation}), which is built up on by \cite{Wolf2018} to \textit{"'combine low level attributes with high level relational knowledge in a generic way"'}. They explain several advantages of including semantic, relational information about the traffic scene over exchanging plain object lists or occupancy grids. In \cite{Petrich2018}, semantic modeling is picked up again and combined with the idea to use \textbf{Probabilistic Entity Relationship} models for state representation under uncertainty. For none of these three approaches did the authors share a complete meta-model or ontology for traffic scenes.
\par
\bigskip

In summary, the previously mentioned takes on modeling and representation of road scenes constitute great building blocks for a CP system. This work aims to combine some of their conceptions to come up with a holistic way to (1) model a traffic scene by all relevant aspects, (2) represent it in an appropriate format and (3) define in what form to efficiently transfer it over the wire (or wirelessly). 

\section{Cooperative Perception}
\label{sec:related_work:cooperative_perception}
Different approaches to Cooperative Perception systems have already been published in the past, each of them placing a different focus.
\par
\bigskip

An early take in the field of V2X communication and – in the broadest sense – also cooperative perception was presented by \cite{Olaverri-Monreal2010}. Based on VANETs, they propose a \textit{see-through system} that allows a driver to virtually perceive the road in front of a large obstacle, e.g. a truck, blocking her sight. To achieve this, DSRC is used to exchange raw camera images between connected participants. However, this approach is somewhat special in a way that it does define an ADAS rather than a CP-system for driverless cars in the sense of this work.
\par
\bigskip

Secondly, as part of the Ko-PER research project \footnote{\url{http://ko-fas.de/41-0-Ko-PER---Kooperative-Perzeption.html}} the authors of \cite{Rauch2011} propose an architecture based on \textbf{track-to-track fusion}, in which local- and global fusion is performed in separate steps to solve the problem of correlated data. Communication is based on IEEE 802.11p and utilizes CAMs and self-defined CPMs, for which they distinguish between iCPMs (infrastructure) and vCPMs (vehicle), depending on the type of sender. Primary objective of their studies is to get detailed insights about \textbf{latency and transmission range} in CP scenarios. Eventually, the system is evaluated in both a laboratory and a real-world setup involving two cars. While their findings are essential for understanding performance-relevant parameters in V2X setups, the evaluation might have been conducted for more realistic scenarios with many traffic participants in addition.
\par
\bigskip

Complementing their previous work in the field of CP, the authors of \cite{Rauch2012} intensively study methods for \textbf{temporal and spatial alignment} of Cooperative Perception messages. A motion model is used to predict observations to current time on the receiver side. In addition, they compare two types of possible transformations to accurately estimate an object's spatial position as combination of local- and received observations. While \cite{Rauch2011} is concerned with communication aspects, \cite{Rauch2012} mainly addresses fusion. Findings from both works are well qualified to be incorporated into a more comprehensive solution. 
\par
\bigskip

While the above studies focus on techniques and parameters within the process of CP itself, \cite{liu2013motion} investigates the impact of CP on \textbf{motion planning} tasks. It is suggested to use local observations for short-term navigation and combined, global observations for longer-term planning. Therefor, a variation of the RRT* algorithm is used on a continuously re-constructed cost map that additionally includes weights corresponding to the underlying observations' \textbf{reliability of perception (ROP)}. Instead of IEEE 802.11p, their connected vehicles use conventional WiFi (802.11n) for P2P communication. Their custom state model does not include any additional information except the probability (confidence) for the quadruple state $X \in \{\chi_{correct}, \chi_{opposite}, \chi_{offroad}, \chi_{roadobst}\}$ of being driveable or not. As a result of evaluating planning quality with respect to trajectory smoothness and accumulated cost they found CP to be beneficial. Another evidence for the positive effect of CP on path planning is given by the authors' subsequent publication \cite{kim2013cooperative}. This complementary work mainly addresses the problems of \textbf{vehicle identification}, \textbf{delay compensation} and merging \textbf{occupancy grid maps}, which are used as an environment model. In contrast to previously mentioned systems, most of which either use high- or feature level fusion, this work is based on low-level fusion, i.e. raw sensor observations (LiDAR and camera) are exchanged. 
\par
\bigskip

The system design proposed by \cite{Calvo2017} is the most comprehensive one in this collection, as it presents a 3+1 level CP architecture to be used as an \textit{obstacle avoidance framework}. DSRC-based V2V communication with IEEE 802.11p is used to exchange participants' immediate local surrounding and long-range, \textbf{cellular}, LTE-based V2I communication facilitates the exchange of more global observations to be used for high-level tasks like traffic re-routing. In addition, their concept also includes I2I communication between RSUs. In comparison to previously mentioned works, which mostly cover certain CP-related aspect, the present paper focuses on conceptually outlining a \textbf{system architecture}. However, neither an environment model nor a concrete state representation or message format are presented and it is not made clear what kinds of information are exchanged at all. 
\par
\bigskip

\cite{Chen2019} is the most recent publication in the field of CP and relies on low-level fusion of 3D LiDAR point clouds. After motivating their approach and outlining several accompanying challenges, the author present their DSRC-based system design. Since raw sensor data is sent, the design does not include the specification of an environment model. Instead, particular emphasis is laid on (1) the \textbf{fusion of point clouds}, (2)\textbf{ deep-learning-based object detection} and (3) \textbf{optimization of network utilization}, i.a. by determining a certain \textbf{region of interest} (ROI) for every frame. As a consequence of their evaluation, the authors find that the effective sensing area can successfully be expanded with CP to capture previously unknown obstacles. 

\section{Cellular V2X Communication}
\label{sec:related_work:cellular_v2x_communication}
With respect to cellular V2X communication – as opposed to DSRC-based solutions – recent standards are being developed by 3GPP most notably with the publication of \cite{3GPP2019}.

In addition, \cite{QualcommTechnologiesInc.2018} conducted comprehensive experiments to measure the performance of LTE and 5G networks with respect to Cooperative Perception and found 5G to be well suited for these use cases. 

This is complemented by the results of \cite{5GAutomotiveAssociation2016}, whose authors found cellular, 5G-based communication to be superior over DSRC / ITS-G5 for CP tasks.

\section{Conclusion}
\label{sec:related_work:conclusion}
Great research has already been contributed to the fields of traffic scene modeling, environment state representation for autonomous driving, cooperative perception and (cellular-) vehicle-to-everything communication. However, there are several limitations with current approaches, that are discussed in greater detail in \autoref{ch:problem_description}. Most notably, the majority of existing CP solutions in literature rely on DSRC-based VANET topologies and in-vehicle data fusion. However, the rise of high-performance cellular networks allows for entirely new concept and system design patterns. Accordingly, this work builds up on previous findings and investigates new possibilities enabled through technological advances to design a holistic, modern system. 