This chapter introduces to essential topics and concepts in the context of this work and provides the reader with background knowledge required to follow in later chapters.

\section{Autonomous Driving}
\label{sec:background:autonomous_driving}
Academia and established industry leaders in automobile manufacturing are vigorously pushing research on autonomous driving technologies alongside emerging start-up companies, who try to enter the new market. The challenge of self-driving cars is believed to be solved within a few decades with high certainty \cite{Frost&SulivanConsulting2018}, although precise forecasts diverge. However, most experts agree that benefits are enormous. Such include decreased risk of collisions and causalities, higher traffic efficiency and less occupied roads – leading to better environmental sustainability – and enhanced driver comfort. New business models – like robo-taxi- or car-sharing services – are likely to arise as transportation culture will undergo a shift from individually owned cars towards a sharing economy and \textit{Mobility as a Service} concepts. Nonetheless, despite these advantages, AD is also accompanied by a number of challenges. Most importantly, government regulations and an appropriate legal framework are vital. Moreover, people are commonly concerned about the accompanying loss of jobs and the cultural changes in general \cite{schoettle2014survey}. 

\subsection{Levels of Autonomy}
\label{subsec:background:labels_of_autonomy}
With reference to self-driving cars, a distinction is usually made between five different levels of autonomy \cite{Klein}. These levels are used to commonly describe vehicles' capabilities and their degree of independence from a human driver with regard to the task of driving. 

\begin{samepage}
\begin{itemize}
	\item \textbf{Level 0:} No computer assistance of any kind. A car is completely controlled by its human driver.
	\item \textbf{Level 1:} Basic assistance, e.g. adaptive cruise control. While most functions are controlled by the driver, the car might take responsibility of a single task, e.g. accelerating and decelerating in certain scenarios.
	\item \textbf{Level 2:} Partial automation, e.g. cruise control and lane centering. At this level, a car is able to take over multiple driving tasks in combination. While the driver is still required to monitor the roadway, she is "'disengaged from physically operating the vehicle"' \cite{Klein} and may keep her hands of the steering wheel and feet of the pedals. To ensure that a driver is still pays full attention and is able to intervene in case of system failures or critical situations, various methods of \textit{Driver Monitoring} are employed. Such include to visually observe a driver's face using cameras or to measure the force applied to the steering wheel.
	\item \textbf{Level 3:} High degree of automation. At this level, a driver might fully rely on a car's self-driving under most conditions, delegating all safety-critical function to the ADAS. Usually, it would maintain a comprehensive awareness of its environment and is able to react on it. Although a driver still has to be present and prepared to take occasional control, she is not required to constantly monitor the traffic. 
	\item \textbf{Level 4:} Full automation. This refers to a system that is able to "'perform all safety-critical driving functions and monitor roadway conditions for an entire trip."' \cite{2016transportation} At this level, there is no necessity for a driver to actually occupy the vehicle.Driving performance of Level 4 autonomous cars is at least equal to human level and would generally even surpass it. 
	\item \textbf{Level 5:} Full autonomy. The highest level of automation describes a system that is capable of driving under any conditions, even extreme ones. Its performance is expected to be at least human-like or even surpass human driving capabilities. 
\end{itemize}
\end{samepage}


With this classification, it is worth noting that only the highest level actually refers to the term "'autonomy"'. \cite{wood2012potential} states that although this term is in more widespread public use, speaking of "'automation"' would be more accurate for levels 1 to 4. Only Level 5 cars are self-governing and may take independent decisions, e.g. selecting a destination and an appropriate route, while cars of all other levels still have a human person in the driver's seat.

\subsection{State of the Art}
\label{subsec:background:state_of_the_art}

\subsection{Sensor Fusion}
\label{subsec:background:sensor_fusion}

\subsection{Autonomous Driving Pipeline}
\label{subsec:background:autonomous_driving_pipeline}
